---
title: "SRA 2019-based"
author: "Julia Levine, Chad Hazlett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(glmnet)
library(dplyr)
library(data.table)
library(ggplot2)
library(reshape2)
library(knitr)

```

\newpage

# Dataset Documentation

The prepared data-set was built using the following sources:

- [Varieties of Democracy](https://www.v-dem.net/en/data/data-version-9/), which is updated annually. 
- Polity data 
- [World Development Indicators](https://datacatalog.worldbank.org/dataset/world-development-indicators), which are updated quarterly. 
- [Uppsala Conflict Data Program](https://ucdp.uu.se/downloads/), which is updated annually. 
- [State of the World Report](https://earlywarningproject.ushmm.org/), which is updated annually.
- [Coup Data](https://www.jonathanmpowell.com/coup-detat-dataset.html): this website contains the most recent updates from Powell and Thyne
- CIA Factbook, which is updated annually. 

The variables taken from each data-set and notes on missingness are discussed below.

## Varieties of Democracy Data

### Variables

- v2elrstrct - candidate restriction
- v2psparban_ord - party ban
- v2psoppaut_ord - barriers to parties
- v2jureform_ord - judicialreform
- v2clrelig_ord - religiousfreedom
- v2xcl_disc - freediscussion
- v2clkill_ord - political killing approved
- v2cldmovem_ord - free movement for men
- v2cldmovew_ord - free movement for women
- v2clsocgrp_ord - social inequality
- v2clrgunev_ord - even civil rights
- v2csreprss_ord - repress civil soc
- v2pepwrsoc_ord - social power dist
- v2pepwrsoc_ord - minority rule
- v2pepwrses - SES power dist
- e_migdppc - GDP per capita, missing since 2017.Filled in using the WDI.
- e_migdpgro - GDP per capita growth, missing since 2017. Filled in using the WDI.
- e_mipopula - population size, missing since 2017. Filled in using the WDI.
- e_cow_exports - country exports, missing since 2017. Filled in using the WDI.
- e_cow_imports- country imports, missing since 2017. Filled in using the WDI.

### Notes

The following variables have not been updated since 2016: e_migdppc, e_mgdpgro, e_mipopula, e_cow_exports, and e_cow_imports. These have been filled in using the WDI data. 

## Polity

- The polity2 and durable variables are no longer updated starting in 2019.



## World Development Indicators

### Variables

- NE.TRD.GNFS.ZS - Trade (% of GDP)
- NY.GDP.PCAP.PP.KD - GDP per capita, PPP (constant 2005 intl $)
- NY.GDP.PCAP.KD - GDP per capita (constant 2000 US$)
- NY.GDP.MKTP.KD.ZG - GDP growth (annual %)
- FP.CPI.TOTL.ZG - Inflation, consumer prices (annual %)
- SP.POP.TOTL- Population, total
- SP.DYN.IMRT.IN - Infant mortality rate

 

## Uppsala Conflict Data Program

### Variables

- battle-deaths

### Notes

- In previous years, we only took the last battle-death observation in a year for each country. In these updates, we sum the total number of battle-deaths each year for each country.
- We also take conflict types 3 and 4 (internal and internationalized internal), where in previous years, we also included conflict types 2 (interstate) from UCDP. 

## State of the World Report

This annual report published on the EWP website is used to update data on mass killings, state-led, and non state-led. 

### Variables

- anymk.start.1
- anymk.start.2window
- anymk.ongoing
- anymk.ever



## Coup Data

### Variables

- coup.try.5yr - indicator variable for whether a coup was attempted within the last five years

### Notes

Gabon and Sudan both had coups in 2019.

## CIA Factbook

The CIA Factbook is used to fill in data on population size, trade-share, and GDP per capita growth if its estimates are more recent than the last present value from the WDI data.

\newpage

The Model 

```{r}

outcomenames <- c("anymk.start.1", "anymk.start.2window")

# Note change to anymk.ongoing and anymk.ever (from mkl.ongoing and mkl.ever)
predictornames <- c("anymk.ongoing","anymk.ever",
                    "reg.afr", "reg.eap", "reg.eur", "reg.mna", "reg.sca", 
                    "countryage.ln", "popsize.ln.combined", "imr.sqrt", 
                    "gdppcgrowth.combined", "ios.iccpr1","includesnonstate",
                    "minorityrule", "elf.ethnic", "battledeaths.ln",
                    "candidaterestriction", "partyban","judicialreform",
                    "religiousfreedom", "pol_killing_approved",
                    "freemove_men4","freemove_women4", "freediscussion",
                    "social_inequality","even_civilrights","repress_civilsoc",
                    "social_power_dist", "ses_power_dist","tradeshare.ln.combined",
                    "coup.try.5yr")
```



```{r}
source("functions_used/create_datasets.R")
source("functions_used/run_model.R")
source("functions_used/format_results.R")

```




## Running the model using 2019 data

```{r, eval = FALSE}
dat <- fread("../Make data/output/prepared2019predictors-2020-08-10.csv")


# run the model and save the output

base_years <- c("2018", "2019")
results <- lapply(base_years, 
                  function(x) format_results(dat = dat, base_year = x))
names(results) <- paste0("base", base_years)
list2env(results, .GlobalEnv)


for(i in 1:length(results)){
  write.csv(results[[i]][1], paste("results/", names(results)[i], "-run-", 
                       Sys.Date(),".csv" , sep = ""), 
            row.names = FALSE)
}

for(i in 1:length(results)){
  write.csv(results[[i]][2], paste("results/", names(results)[i], "-coeffs-run-", 
                       Sys.Date(),".csv" , sep = ""), 
            row.names = FALSE)
}

for(i in 1:length(results)){
    cv.glmnet <- results[[i]][[3]]
  save(cv.glmnet, predictornames, file=paste("results/", names(results)[i], "-cv.glmnet-", 
                       Sys.Date(),".Rdata" , sep = ""))
}


```


