---
title: "Updating Data to be used in SRA"
author: "Julia Levine"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(knitr)
library(zoo)
library(ggplot2)
# the year of the data we are adding in
update_years <- c(2019)
first.pass <- F
```

\newpage




# Load in old data

The following code reads in the old data. It then takes the most recent data for each country and replicates it for the years that we are updating -- this let's us easily keep the time-invariant values, like region. I then set aside the old data, and update the data that will be appended. 

```{r}
# read in data from the previous years , called dat
  
  load("input/prepared2018predictors_2_11_20.RData")
  dat[, last_year := NULL]

# carry forward all data from the most recent year

  carry_from <- max(dat$year)
  
  # this mainly applies to updating multiple years simultaneously
  # it carries forward the most recent data for each of the years to be updated
  dat_new <- rbindlist(rep(list(dat[year == carry_from]), times = length(update_years)))

  # relabel the years
  
  dat_new$year <- rep(update_years, each = nrow(dat_new))


# rename old data and leave it alone until you've updated all 
# variables except for the lag/leads
  
  dat_old <- dat
  rm(dat)

```



# Update 

```{r}
# function to carry forward values of a variable for countries where it's NA

  carry_forward <- function(variable, carry_to = 2019, data = dat, carry_from = 2018){
    na_var <- data[year == carry_to & is.na(get(variable))]$sftgcode
    
    cat(paste0("Missing values for ", length(na_var), " countries."))
    
    carry_var <- subset(data[year == carry_from & sftgcode %in% na_var], 
                        select = c("sftgcode", variable))
    cat(paste0("\nFilling in values for ", 
               nrow(carry_var[!is.na(get(variable))]), " countries."))

    for(i in 1:length(na_var)){
      data[year == carry_to & sftgcode == na_var[i], 
          (variable) := carry_var[sftgcode == na_var[i], get(variable)]]
    }
  }
  
```



## Country age

This chunk updates country age by adding the difference between the updated years and the year the data was carried from to the country age variable. 

```{r}
# update country age

  diff <- update_years - carry_from
  
  for(i in 1:length(update_years)){
    dat_new$countryage[dat_new$year == update_years[i]] <- 
    dat_new$countryage[dat_new$year == update_years[i]] + diff[i]
  }

# update logged country age variable

  dat_new$countryage.ln <- log(dat_new$countryage + 1)
```

## merge_dat function

The following is a function that removes the variables that were carried forward from the base data, and then merges in the updated variables from the appropriate year.

```{r}
merge_dat <- function(new_data, base_data = dat_new, 
                      merge_new_by = c("sftgcode", "year"), 
                      merge_base_by = c("sftgcode", "year")){
  new_data_vars <- colnames(new_data)[!colnames(new_data) %in% merge_new_by]
  dat1 <- subset(base_data, 
                 select = colnames(base_data)[!(colnames(base_data) %in% 
                                                  new_data_vars)])
  dat1 <- merge(dat1, new_data, 
                by.x = merge_base_by, 
                by.y = merge_new_by, all.x = TRUE)
  dat1
}
```

## VDEM

The following code reads in the most recent V-DEM data, selecting the relevant variables and renaming them. 

```{r, eval = FALSE}
# read in data

  vdem <- fread("input/Country_Year_V-Dem_Full+others_CSV_v10/V-Dem-CY-Full+Others-v10.csv")

# subset to just use these variables

  vdem_vars <- c("v2cldmovem_ord", 
                 "v2cldmovew_ord", 
                 "v2clkill_ord",
                 "v2clsocgrp_ord", 
                 "v2clrgunev_ord", 
                 "v2csreprss_ord",
                 "v2pepwrsoc_ord", 
                 "v2elrstrct", 
                 "v2psparban_ord",
                 "v2psoppaut_ord", 
                 "v2jureform_ord", 
                 "v2clrelig_ord",
                 "v2xcl_disc", 
                 "v2pepwrses_ord", 
                 "v2pepwrses", 
                 "e_migdpgro",
                 "e_mipopula", 
                 "e_cow_exports", 
                 "e_cow_imports", 
                 "e_migdppc"
                 )
  keep <- c("COWcode", "country_name", "year", vdem_vars)

# function that renames and formats variables, 
# takes dataset and vector of variables to keep as inputs
  
  source("helper scripts/format_vdem.R")
  
# function that maps COW codes onto PITF country codes, 
# contained in format_vdem 
  
  source("helper scripts/cowtopitf2018.R")

# format and save vdem
  
  vdem <- format_vdem(dat = vdem, keep = keep)
  save(vdem, file = "input/temp/vdem.Rdata")
```




This chunk removes the old year's V-DEM variables, summarizes missingness, and merges in the new ones.

```{r}
# read in formatted VDEM variables for the update years 
  
  load("input/temp/vdem.Rdata")

# limit to the update years
  
  vdem <- vdem[year %in% update_years]

# count missing values
# check which variables are missing values
  
  missing <- apply(vdem, 2, function(x) sum(is.na(x)))
  kable(missing, col.names = "NA count")

# check which variables are missing for all countries, indicating that they
# have not been updated yet, and are potentially discontinued
  
  all.missing <- names(missing)[missing == nrow(vdem)]
  cat(paste0("The following variables have not been updated as of ", 
            Sys.Date(),":\n", paste(all.missing, collapse = "\n")))
  
  
# this ensures that we dont replace old country names
# VDEM sometimes changes them
  
  vdem$country_name <- NULL

# update V-DEM data
  
  dat_new <- merge_dat(new_data = vdem)
  dat_new[, tradeshare.ln := log(tradeshare)]  
```





## Polity 

We've decided to drop the polity variables, but I'm just updating the durable and pol.durable variables here the same way I updated country age.

```{r}
# update durable the same way we update country age
# Note: pol.durable and durable are the same thing

  diff <- update_years - carry_from
  
  for(i in 1:length(update_years)){
    dat_new$durable[dat_new$year == update_years[i]] <-
    dat_new$durable[dat_new$year == update_years[i]] + diff[i]
  
    dat_new$pol.durable[dat_new$year == update_years[i]] <-
    dat_new$pol.durable[dat_new$year == update_years[i]] + diff[i]
  }

# update the logged variables

  dat_new$pol.durable.ln <- log(dat_new$pol.durable + 1)
  dat_new$durable.ln <- dat_new$pol.durable.ln 
```







## WDI

This chunk pulls select variables from the WDI API.



```{r, eval = FALSE}
# In June 2020 WDI decommissioned V1 of the API
# so you may need to re-install the WDI package.  

  # devtools::install_github("vincentarelbundock/WDI")
  library(WDI)

# list indicators we want to pull
  
  wdilist <- c("NE.TRD.GNFS.ZS",     # Trade (% of GDP)
               "NY.GDP.PCAP.PP.KD",  # GDP per capita, PPP (constant 2005 intl $)
               "NY.GDP.PCAP.KD",     # GDP per capita (constant 2000 US$)
               "NY.GDP.MKTP.KD.ZG",  # GDP growth (annual %)
               "FP.CPI.TOTL.ZG",     # Inflation, consumer prices (annual %)
               "SP.POP.TOTL",        # Population, total
               "SP.DYN.IMRT.IN"      # Infant mortality rate
               #"NE.EXP.GNFS.ZS",     # exports (% of GDP)
               #"NE.IMP.GNFS.ZS"      # imports % of GDP
               
      )

  # confirm that all of the old indicator names are in use 
  
    (check.names <- rbindlist(lapply(wdilist, function(x) {
      # search the indicator list for each of the variables in wdilist
       res <- WDIsearch(x, field = "indicator") 
       # take the first match (drops those with extra letters on the end)
       if(length(res) > 2){
         res <- res[1, ]
       } 
       data.table("indicator" = res[1], "name" = res[2])})))

# Extract latest version of desired variables from WDI 
# also pull 5 years from before so that we can carry forward
# more recent values for variables like infant mortality that are slow to update
  
  wdi <- WDI(country="all", indicator=wdilist, extra=FALSE, 
             start=(min(update_years) - 5))

# Add PITF country codes for merging 
  source("helper scripts/f.pitfcodeit.R")
  wdi <- pitfcodeit(wdi, "country")
  wdi$country <- as.character(wdi$country)

# Subset to drop cases with missing PITF codes, cut extra id vars

  wdi <- subset(wdi, !is.na(sftgcode), select=-c(1, 2))

# Reorder for easier review
  
  wdi <- wdi[order(wdi$sftgcode, wdi$year),]

# Rename variables-- add a "new" to indicate these are newly brought in from wdi 
# to avoid reusing names already in the old EWP data
  
  names(wdi) <- c("year",
                  "wdi.trade.new",
                  "wdi.gdppcppp.new",
                  "wdi.gdppc.new",
                  "wdi.gdppcgrow.new",
                  "wdi.inflation.new",
                  "wdi.popsize.new", 
                  "wdi.imrate.new", 
                  #"wdi.expshare.new", 
                  #"wdi.impshare.new",
                  "sftgcode"
                  )  

# save to the input folder
  
  fwrite(wdi, paste0("input/temp/wdi/pulled-", Sys.Date(), ".csv"))
```




This chunk reads in the latest pull from the WDI API, counts the missing values, and merges it into the main dataset.

```{r}
# list all files in the wdi directory and read in the latest pull

  wdi.files <- list.files("input/temp/wdi/")
  latest.pull <- which.max(lapply(wdi.files, 
                                  function(x) {
                                    as.Date(gsub("pulled-|.csv", "", x))}))
  wdi <- fread(paste0("input/temp/wdi/", wdi.files[latest.pull]))
  new.wdi <- wdi[year %in% update_years]
  
# check which variables are missing values
  
  missing <- apply(subset(new.wdi, 
                          select = -c(1, ncol(new.wdi))), 2, 
                   function(x) sum(is.na(x)))
  kable(missing, col.names = "NA count")

# check which variables are missing for all countries, indicating that they
# have not been updated yet, but potentially will be in the future
  
  all.missing <- names(missing)[missing == length(unique(wdi$sftgcode))]
  cat(paste(all.missing, "not updated as of", Sys.Date()))
  
  # we have been carrying forward a 2015 version of imrate
  # update the value to be carried forward to the most recent non-missing value
  
  old.wdi <- wdi[year < min(update_years) & !is.na(wdi.imrate.new)]
  to.carry <- old.wdi[, .("wdi.imrate.new" = wdi.imrate.new[which.max(year)]), 
                      by = "sftgcode"]
  to.carry[, year := update_years]
  
  new.wdi <- merge_dat(new_data = to.carry, base_data = new.wdi)
  
  # do the same thing for missing values of tradeshare and gdp as they were 
  # mostly updated in 2018 and the earliest values from CIA factbook are  2017

  wdi <- rbind(new.wdi, old.wdi)
  
  carry_forward(variable = "wdi.trade.new", carry_from = 2018, data = wdi)
  carry_forward(variable = "wdi.trade.new", carry_from = 2017, data = wdi)
  carry_forward(variable = "wdi.trade.new", carry_from = 2016, data = wdi)
  
  
  carry_forward(variable = "wdi.gdppc.new", carry_from = 2018, data = wdi)
  carry_forward(variable = "wdi.gdppc.new", carry_from = 2017, data = wdi)
  carry_forward(variable = "wdi.gdppc.new", carry_from = 2016, data = wdi)

  
  missing <- apply(subset(wdi[year %in% update_years], 
                          select = -c(1,2)), 2, 
                   function(x) sum(is.na(x)))
  kable(missing, col.names = "NA count")
  
# Merge it in to main data
  
  dat_new <- merge_dat(new_data = wdi[year %in% update_years])
  dat_new[, imr.sqrt := sqrt(wdi.imrate.new)]
  dat_new[, wdi.trade.ln.new := log(wdi.trade.new)]
  
```






## UCDP battle-related deaths 

The following code reads in the UCDP data and changes country names to match. 
Below, I sum the number of battledeaths for each country in the update years, and merge this onto dat_new. 

```{r, eval = F}
# read in UCDP data
  
  ucdp <- fread("input/ucdp-brd-dyadic-201.csv")

# making sure we have the same variables as in 18.1 version. 
# colnames for 18.1 are uppercase, use tolower when checking

  colnames(ucdp) <- gsub("_", "", colnames(ucdp))

# limit to the update years and conflict type >= 3
  
  ucdp <- ucdp[year %in% update_years & typeofconflict >= 3]

# change country names to match
  
  diff_loc <- setdiff(unique(ucdp$locationinc), unique(dat_new$country_name))
  
  # drops the locations that involve multiple countries
  
    diff_loc <- diff_loc[grepl(", ", diff_loc) == F]
    diff_loc
  
  # change names
    
    ucdp$locationinc[ucdp$locationinc=="Russia (Soviet Union)"] = "Russia"
    ucdp$locationinc[ucdp$locationinc=="Myanmar (Burma)"] = "Burma/Myanmar"
    ucdp$locationinc[ucdp$locationinc=="Yemen (North Yemen)"] = "Yemen"
    ucdp$locationinc[ucdp$locationinc=="DR Congo (Zaire)"] = "Democratic Republic of Congo"
    
    if(length(setdiff(unique(ucdp$locationinc), 
                      unique(dat_new$country_name)))>0){
      stop("Different country names")}

# group by year/country and sum battledeaths
    
  bd <- ucdp[, .("battledeaths" = sum(bdbest)), by = c("locationinc", "year")]
  
# save battledeath data
  
  fwrite(bd, "input/temp/battledeaths.csv")
```

I then read in the battledeaths for each country and merge it into the new data, replacing missing values with zeroes and updating the logged version of this variable.

```{r}
# read in battledeaths data

  bd <- fread("input/temp/battledeaths.csv")

# merge battledeaths into main data
    
  dat_new <- merge_dat(new_dat = bd,
                       merge_new_by = c("locationinc", "year"), 
                       merge_base_by = c("country_name", "year"))
    
  
# fill in zeroes and update logged variable
  
  dat_new[, battledeaths := ifelse(is.na(battledeaths), 0, battledeaths)]
  dat_new[, battledeaths.ln := log(battledeaths + 1)]
```


## Coup Attempts

This code reads in the most updated version of the Powell and Thyne data which is posted on their website. It selects coups in the years to be updated

```{r, eval = FALSE}
# pull coup data from powell and thyne websit

  coup_dat <- as.data.table(read.delim("http://www.uky.edu/~clthyn2/coup_data/powell_thyne_coups_final.txt"))
  coup_dat[year == 2016]
# keep coups in the update years
  
  new_coup <- coup_dat[year %in% update_years]

# where coup == 2 it was a successful coup
# where coup == 1 it was a failed coup
  
  new_coup[, cou.s.d := ifelse(coup == 2, 1, 0)]
  new_coup[, cou.f.d := ifelse(coup == 1, 1, 0)]

# save coup data
  
  fwrite(new_coup, paste0("input/temp/", "coups-pulled-", Sys.Date(), ".csv"))
```


```{r}
inscr.coup <- readxl::read_excel("input/CSPCoupsAnnualv2018.xls")
setDT(inscr.coup)[year >= 2016 & atcoup2 == 1]
```


This updates the coup variables in dat_new.

```{r}
# read in the coup data

  temp.files <- list.files("input/temp")
  new_coup <- fread(paste0("input/temp/", temp.files[grep("coups", temp.files)]))

# select variables and print
  
  new_coup <- subset(new_coup, 
                     select = c("country", "year", "cou.s.d", "cou.f.d"))
  kable(new_coup)
  
# merge coup data into dat_new
  
  dat_new <- merge_dat(new_data = new_coup, 
                       merge_new_by = c("country", "year"), 
                       merge_base_by = c("country_name", "year"))

# fill in missing values with zeroes

  coup_cols <- c("cou.s.d", "cou.f.d")
  dat_new[, (coup_cols) := lapply(.SD, function(x) 
    ifelse(is.na(x), 0, x)), .SD = coup_cols]

# update cou.any variable (1 if there was either a failed or successful coup)
  
  dat_new[, cou.any := ifelse(cou.s.d>0 | cou.f.d>0, 1, 0)]

```


## Mass Killing Variables 

There were no starts or ends to mass killing events in 2019.

```{r}
# initialize state led mk vars

  dat_new$mkl.end <- 0
  dat_new$mkl.start <- 0
  
# initialize non-state led mk vars
  
  dat_new$nonstatemk.end <- 0
  dat_new$nonstatemk.start <- 0
```



# Append new data

This appends dat_new to dat_old, creating the full data-set

```{r}
# check to make sure you can append the new data

  if(length(setdiff(colnames(dat_old), colnames(dat_new)))>0){
    stop("different colnames")}
  if(length(setdiff(colnames(dat_new), colnames(dat_old)))>0){
    stop("different colnames")}

# append new data
  
  setcolorder(dat_new, colnames(dat_old))
  dat <- rbind(dat_old, dat_new)
```



## Mass Killing Onsets

I now make changes to the mass killing variables in years prior to the update years. 

There is one change to be made: the end of the non-state led mass killing in Thailand in 2016. 


```{r}
# change old data
  
  dat[, nonstatemk.end := ifelse(sftgcode == "THI" & year == 2016, 1, nonstatemk.end)]
  dat[, nonstatemk.ongoing := 
        ifelse(sftgcode == "THI" & year > 2016, 0, nonstatemk.ongoing)]

# combined onset variables
  
  dat[, anymk.start := ifelse((!is.na(mkl.start) & mkl.start == 1)|
                                nonstatemk.start == 1, 1, 0)] 
  
  dat[, anymk.ongoing := ifelse((!is.na(mkl.ongoing) & mkl.ongoing == 1)|
                                nonstatemk.ongoing == 1, 1, 0)] 

  dat[, anymk.ever := ifelse((!is.na(mkl.ever) & mkl.ever == 1)|
                                nonstatemk.ever == 1, 1, 0)] 

```



```{r, eval = F}
# save mass killing variables

  mkl_dat <- subset(dat, select = c("country_name", "sftgcode", "year", 
                                    "mkl.start", "mkl.end", "mkl.ongoing", 
                                    "mkl.ever", "nonstatemk.start", 
                                    "nonstatemk.end", "nonstatemk.ongoing", 
                                    "nonstatemk.ever"))
  setkey(mkl_dat, country_name, year)

fwrite(mkl_dat, file = "mkl_data.csv", row.names = F)
```





# Leads and Lags

In this section, I update the variables that involve leads and lags

## MK lead variables

```{r}
# read in list of the first time Sftgcodes are used in PITF data

  map <- fread("../../../EWP (1)/2019SRA/Make data/sftg_name_map.csv")
  setnames(map, "V1", "min_ewp_year")
  
  # aggregate over different country names and drop duplicates
    
    map <- map[country != ""]
    map[, min_ewp_year := min(min_ewp_year), by = "sftgcode"]
    map[, ':=' (country = NULL, country_name = NULL)]
    map <- map[!duplicated(map)]


# merge in first year a sftgcode was used in the PITF data
    
  dat <- merge(dat, map, by = "sftgcode", all.x = T)

# order by year, country name, and sftgcode
    
  setkey(dat, year, country_name, sftgcode)

# create anymk lead variable, shifting by sftgcode
    
  dat[, anymk.start.1 := shift(anymk.start, 1, type = "lead"), by = "sftgcode"]
    
  # For new states we don't attribute mass killings to their origin country.
  # if the year is the year before the PITF starts first used an sftgcode, 
  # then reset to 0
    
    dat[, anymk.start.1 := ifelse(year == min_ewp_year - 1, 0, anymk.start.1)]
    
    
#  create the rest of the lead variables based off of the one-year lead
    
  dat[, anymk.start.2 := shift(anymk.start.1, type = "lead"), by = "sftgcode"]
  dat[, anymk.start.3 := shift(anymk.start.1, n= 2, type = "lead"), by = "sftgcode"]
  
  dat[, anymk.start.2window := as.double((anymk.start.1 + anymk.start.2) > 0) ]
  dat[, anymk.start.3window := as.double((anymk.start.2window + anymk.start.3) > 0)]
```





## Coup in the last 5 years

Create dataframe of sftgcodes that are created as a result of a coup. We then turn the coup.try.5yr indicator to 1 for the first four years of that sftgcode's existence.

```{r}
# this is a table of sftgcodes created by a coup

  new_sftg <- data.table("sftgcode" = c("GAB", "ETI", "DJI"), 
                         "coup.create.yr" = c(1964, 1989, 2000))

# merge this into main

  dat <- merge(dat, new_sftg, by = "sftgcode", all = T)

# create a variable for the last year when an sftgcode had a coup
  
  dat[, last_coup_yr := ifelse(cou.any == 1, year, NA)]
  dat[, last_coup_yr := na.locf(last_coup_yr, fromLast = F, na.rm = F), 
      by = "sftgcode"]
  dat[, last_coup_yr := ifelse(is.na(last_coup_yr), -Inf, last_coup_yr)]

# for sftgcodes that were initiated due to a coup, 
# replace -Inf with the year of the coup
  
  dat[, last_coup_yr := ifelse(!is.na(coup.create.yr) & 
                                 coup.create.yr > last_coup_yr, 
                               coup.create.yr, last_coup_yr)]
  
# create variable for whether there was a coup in the last 5 years
  
  dat[, coup.try.5yr := as.double((year - last_coup_yr) <= 4)]
```





# Data Preparation


## Remake some variables

This section fills in some V-DEM variables by creating an adjusted version of the relevant WDI variable. This significantly cuts down on missingness.

```{r}
make.combined <- function(vdem.var, wdi.proxy){
  # Make an adjusted version of the wdi one to fit the tradeshare (VDEM) one:  
    # drop update year as we've carried forward some values to this year
  
    lm.adjust <- lm(as.formula(paste0(vdem.var, "~", wdi.proxy)), 
                    data = dat[year < update_years])
    
  # fill in update years with fitted values from regression on all years  
    
    dat[year %in% update_years, lm.adjust := coef(lm.adjust)[1] + 
          coef(lm.adjust)[2]*get(wdi.proxy)]
  
  # Where vdem var is missing, replace with the adjusted wdi proxy
    combined.var <- paste0(vdem.var, ".combined")
    dat[year %in% update_years, 
        (combined.var) := 
          ifelse(!is.na(get(vdem.var)), get(vdem.var), get(wdi.proxy))]
    
  # summarize differnce in missingness
    na.old <- sum(is.na(subset(dat[year %in% update_years], select = vdem.var)))
    na.new <- sum(is.na(subset(dat[year %in% update_years], select = combined.var))) 
  
    cat(paste0("In the update years we were missing ", na.old, 
           " observations for ", vdem.var))
    cat(paste0("\nUsing the WD indicator, ",  wdi.proxy, 
               ", as a proxy we are missing ",  na.new, " observations"))

    dat[, lm.adjust := NULL]
}    
```



```{r}
# create combined variable for tradeshare

  make.combined(vdem.var = "tradeshare.ln", wdi.proxy = "wdi.trade.ln.new")

# create combined variable for popsize

  make.combined(vdem.var = "popsize", wdi.proxy = "wdi.popsize.new")
  
  # update logged version
  
    dat[year %in% update_years, popsize.ln.combined := log(popsize.combined)]

# create combined variable for GDP per capita growth

  make.combined(vdem.var = "gdppcgrowth", wdi.proxy = "wdi.gdppcgrow.new")

```



# Missingness

This section looks at remaining missingness and attempts to fill in values. 

```{r}
# select variables we would like to check missingness for

  predictornames <- c("anymk.ongoing","anymk.ever",
                      "reg.afr", "reg.eap", "reg.eur", "reg.mna", "reg.sca", 
                      "countryage.ln", "popsize.ln.combined", "imr.sqrt",
                      "gdppcgrowth.combined", "ios.iccpr1","includesnonstate",
                      "durable.ln","minorityrule", "elf.ethnic", 
                      "battledeaths.ln", "candidaterestriction", "partyban", 
                      "judicialreform", "religiousfreedom", 
                      "pol_killing_approved", "freemove_men4", 
                      "freemove_women4", "freediscussion",
                      "social_inequality","even_civilrights","repress_civilsoc", 
                      "social_power_dist", "ses_power_dist",
                      "tradeshare.ln.combined", 
                      "coup.try.5yr", "polity2.fl.2","polity2.fl.3")

# look for missingness in the update years for select variables

  dat.check <- subset(dat, year %in% update_years, 
                      select = c("country_name","year", predictornames))

# for each of the update years, count the number of NAs for each variable
# select variables with positive NA counts

  comp <- lapply(update_years, 
                 function(y) apply(dat.check[year == y], 
                                   2, function(x) sum(is.na(x))))
  na.count <- unlist(comp)
  years <- rep(update_years, each = ncol(dat.check))
  check <- data.table(na.count, years, "var"=names(na.count))
  check <- check[na.count > 0]

# visualize missingness
  
  (missing1 <- ggplot(check) + 
      geom_bar(aes(y = na.count, x = var, fill = factor(years)), 
               stat = "identity")+ 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      labs(x = "Missing Variables", y = "Number of Observations Missing", 
           fill = "Year", 
           title = "Number of Countries with Missing Values Before Fixing"))

```




## Fix missingness

This chunk defines functions to look at patterns in NAs and carry forward values.

```{r}

# function to look at patterns in NAs over time for each country with missing values

  look_na <- function(variable, data = dat){
    missing <- data[is.na(get(variable)) & 
                     year %in% update_years, c("year", "country_name")]
    countries <- unique(missing$country_name)
    if(length(countries) == 0){
      look <- "No missing values"
    }else{
      look <-lapply(1:length(countries), function(x){
        out <- subset(data, country_name == countries[x],
                      select = c("year", variable)) 
        colnames(out) <- c("year", countries[x])
        out})
      
      # merge all data.tables in the list by year
        look <- Reduce(function(...) merge(..., all = T, by = "year"), look)
        colnames(look) <- c("year", as.character(countries))
        look <- look[order(look$year, decreasing = TRUE), ]
      }
    look
  }


# create blank spreadsheets to fill out
  
  fill.tab <- function(var){
    out <- subset(dat[is.na(get(var)) & year %in% update_years], 
                  select = c("country_name", var))
    out[, cia.factbook.est := NA]
    out[, est.year := NA]
    fwrite(out, file = paste0("input/temp/missing/", var, ".csv"))
  }
  
  
# look at last present value for countries  missing data in update years
  
  last.present <- function(var){
    look.na.long <- melt(look.na.vars[[var]], id.vars = "year", 
                         variable.name = "country_name")
    out <- look.na.long[!is.na(value), .("carried_value" = value[which.max(year)], 
                                  "carry_from" = max(year)), by = "country_name"]
    out
  }
```




```{r}
# for each of the variables that we have countries with missing values
# look at the values in the last few years for that country

  look.na.vars <- lapply(c(check$var, "wdi.gdppc.new"), look_na)
  names(look.na.vars) <- c(check$var, "wdi.gdppc.new")


  if(first.pass == T){
    lapply(check$var, fill.tab)
    fill.tab("wdi.gdppc.new")
  }  
```

### Candidate Restriction

```{r}      
# candidate restriction

  head(look.na.vars$candidaterestriction)
  
# just carry forward this value 
  
  carry_forward("candidaterestriction")

```

For the variables addressed below, I read in values from the CIA Factbook, and carry those values forward if they're more recent than the last present variable, with several exceptions. Infant mortality rate and population size variable show no change despite the estimates being from more recent years. Recall that I carried forward 2018 infant mortality rates from the WDI already. 

GDP per capita growth and tradeshare are carried forward from the CIA Factbook if the estimate is from a year after 2015, as we had been carrying forward these values from 2015 previously. 

In the tables presented for each variable, I show the value that is filled in for that country in column 2. This is either the "carried_value", carried from the year listed under "carry_from", or it is a transformed version of the value under "cia.factbook.est".

### Infant Mortality

```{r}

  # head(look.na.vars$imr.sqrt)

# read in CIA Factbook estimates

  cia.factbook <- fread("input/temp/missing/imr.sqrt.csv")
  
  comp <- merge(cia.factbook, last.present("imr.sqrt"), 
                by = "country_name")
  comp[, imr.sqrt := ifelse(carry_from > est.year, 
                                        carried_value, sqrt(cia.factbook.est))]
  kable(comp)
  
# replace missing

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country_name ==  country.fill$country_name & year == update_years, 
        imr.sqrt := country.fill$imr.sqrt]
  }

```




### Population size


```{r}

  # head(look.na.vars$popsize.ln.combined)


# read in CIA Factbook estimates

  cia.factbook <- fread("input/temp/missing/popsize.ln.combined.csv")
  
  comp <- merge(cia.factbook, last.present("popsize.ln.combined"), 
                by = "country_name")
  comp[, popsize.ln.combined := ifelse(carry_from > est.year, 
                                        carried_value, log(cia.factbook.est))]
  comp[, popsize.combined := ifelse(carry_from > est.year, 
                                        carried_value, cia.factbook.est)]
  kable(comp)
  
# replace missing

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country_name ==  country.fill$country_name & year == update_years, 
        popsize.ln.combined := country.fill$popsize.ln.combined]
  }

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country_name ==  country.fill$country_name & year == update_years, 
        popsize.combined := country.fill$popsize.combined]
  }
```


### GDP Per capita growth


```{r}

# read in CIA Factbook estimates

  cia.factbook <- fread("input/temp/missing/gdppcgrowth.combined.csv")
  comp <- merge(cia.factbook, last.present("gdppcgrowth.combined"), 
                by = "country_name")
  # these values were carried to 2018 from 2016 or 2017 so let's overwrite them
  comp[, gdppcgrowth.combined := ifelse(est.year < 2016, 
                                        carried_value, cia.factbook.est/100)]
  kable(comp, digits = 3)
  
# replace missing

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country_name ==  country.fill$country_name & year == update_years, 
        gdppcgrowth.combined := country.fill$gdppcgrowth.combined]
  }

```


### tradeshare.ln.combined

#### First have to fill in GDP per capita

```{r}

  cia.factbook <- fread("input/temp/missing/wdi.gdppc.new.csv")
  comp <- merge(cia.factbook, last.present("wdi.gdppc.new"), 
                by = "country_name", all.x = T)
  comp[, wdi.gdppc.new := ifelse((est.year < carry_from & !is.na(carry_from))|
                                   is.na(cia.factbook.est), carried_value, 
                                 cia.factbook.est)]
  kable(comp, digits = 3)
  
  # replace missing

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country_name ==  country.fill$country_name & year == update_years, 
        wdi.gdppc.new := country.fill$wdi.gdppc.new]
  }
```



```{r}
# read in CIA Factbook estimates

  cia.factbook <- fread("input/temp/missing/tradeshare.ln.combined.csv")
  comp <- merge(cia.factbook, last.present("tradeshare.ln.combined"), 
                by = "country_name")
  other.vars <- dat[year == update_years, 
                    c("wdi.gdppc.new", "popsize.combined", "country_name")]
  comp <- merge(comp, other.vars, by= "country_name", all.x = T)
  
  # these values were carried to 2018 from 2016 or 2017 so let's overwrite them
  # as long as factbook estimates are at least from 2016
  
  # The VDEM imports and exports are in millions of USDs, then x10^6. 
  # tradeshare = (imports + exports (in trillions))/(gdppcgrowth + popsize)
  # The CIA factbook data is in billions, so needs to be multiplied by 1e+9
  
    comp[, tradeshare.ln.combined := 
           ifelse(est.year < 2016, carried_value, 
                  log(1e+9*cia.factbook.est/(wdi.gdppc.new*popsize.combined)))]
    kable(subset(comp, select = c("country_name", "tradeshare.ln.combined",
                                  "cia.factbook.est", "est.year", 
                                  "carried_value", "carry_from")))

# replace missing

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country_name ==  country.fill$country_name & year == update_years, 
        tradeshare.ln.combined := country.fill$tradeshare.ln.combined]
  }
    
# carry forward for Somalia
  carry_forward("tradeshare.ln.combined")
  
```



## Check that all is filled in

```{r}
# repeat to make sure there are no more missing
    
  look.na.vars <- lapply(check$var, function(x) 
    look_na(x, dat = dat[country_name!="Taiwan"]))
  names(look.na.vars) <- check$var
  look.na.vars
  
```



```{r}
fwrite(dat, file = paste0("output/prepared2019predictors-", Sys.Date(), ".csv"))
```

# Checking the append

```{r}
# function to check all.equal for each var

  compare_var <- function(var){
    vars <- paste(var, c("x", "y"), sep = ".")
    
    check <- subset(compare, select = c("year", "sftgcode", vars[1], vars[2]) )
    colnames(check) <- c("year", "sftgcode", "var.x", "var.y")
    
    all.equal(check$var.x, check$var.y)
  }


# look into differences where all.equal == F for each var

  check_diffs <- function(var){
    vars <- paste(var, c("x", "y"), sep = ".")
    
    check <- subset(compare, select = c("year", "sftgcode", vars[1], vars[2]) )
    
    check[get(vars[1]) != get(vars[2])|
            is.na(get(vars[1])) & !is.na(get(vars[2]))|
            is.na(get(vars[2])) & !is.na(get(vars[1]))]
  }

```



```{r}
# read in newly updated data and just focus on the last year in old

  dat <- fread("output/prepared2019predictors-2020-08-10.csv")
  new_dat <- dat[year <= 2018]

# read in original base data
  
  load("input/prepared2018predictors_2_11_20.RData")
  old_dat <- dat
  rm(dat)

# check row count
  
  if(nrow(old_dat) != nrow(new_dat)){stop("Different observation count")}

# merge new and old data to compare
  
  compare <- merge(new_dat, old_dat, by = c("sftgcode", "year"))

# make sure they match on all variables in the old data
  
  check.vars <- colnames(old_dat)[!colnames(old_dat) %in% c("sftgcode", "year")]
  
  comp <- data.table("all.equal" = sapply(check.vars, compare_var), 
                     "var" = check.vars)
  
# look into the differences where all.equal == F
# and where the issue isn't an NA value mismatch -- come back to this later
  
  differences <- comp[all.equal!="TRUE"][!grep("is.NA", all.equal)]
  diff.look <- lapply(differences$var, check_diffs)
  names(diff.look) <- differences$var
  
  cat(paste0("There are ", length(diff.look), 
             " variables where the two datasets disagree on non-missing values."))

  for(i in 1:length(diff.look)){
    print(diff.look[[i]])
  }

# now let's look at differences in missing values
  
  na.diffs <- comp[all.equal!="TRUE"][grep("is.NA", all.equal)] 
  na.diff.look <- lapply(na.diffs$var, check_diffs)
  names(na.diff.look) <- na.diffs$var
  
  # for each variable, count the instances of differences in NAs by year
  count.by.yr <- rbindlist(lapply(na.diff.look, 
                                  function(x) x[, .N, by = "year"]))
  count.by.yr[, var := na.diffs$var]
  # We can see that the only differences are the filling of NAs
  # for the appropriate lead variables
  count.by.yr
```

